---
title: "Overiew of R Modelling Packages"
author: "Daniel Lüdecke"
output: html_document
---

```{r setup, include=FALSE}
library(knitr)
library(docxtractr)

knitr::opts_chunk$set(echo = TRUE)

docx <- read_docx("Regressionspakete.docx")
tab <- docx_extract_tbl(docx)
colnames(tab) <- c("Nature of Response", "Example", "Type of Regression", "R package or function")
```

This is an overview of R packages and functions for fitting various different types of regression models. For each row, the upper cells in the last column (packages and functions refer to "simple" models, while the lower cells refer to their mixed models counterpart (if available and known).


```{r echo=FALSE}
kable(tab)
```

Included packages for non-mixed models:

  - Base R: `lm()`, `glm()`
  - [AER](https://CRAN.R-project.org/package=AER): `tobit()`
  - [aod](https://CRAN.R-project.org/package=aod): `betabin()`
  - [betareg](https://CRAN.R-project.org/package=betareg): `betareg()`
  - [brglm2](https://CRAN.R-project.org/package=brglm2): `bracl()`, `brmultinom()`
  - [censReg](https://CRAN.R-project.org/package=censReg): `censReg()`
  - [HRQoL](https://CRAN.R-project.org/package=HRQoL): `HRQoL()`
  - [MASS](https://CRAN.R-project.org/package=MASS): `glm.nb()`, `polr()`
  - [nnet](https://CRAN.R-project.org/package=nnet): `multinom()`
  - [ordinal](https://CRAN.R-project.org/package=ordinal): `clm()`, `clm2()`
  - [pscl](https://CRAN.R-project.org/package=pscl): `zeroinfl()`, `hurdle()`
  - [VGAM](https://CRAN.R-project.org/package=VGAM): `vglm()`

Included packages for mixed models:

  - [glmmTMB](https://CRAN.R-project.org/package=glmmTMB): `glmmTMB()`
  - [lme4](https://CRAN.R-project.org/package=lme4): `lmer()`, `glmer()`, `glmer.nb()`
  - [MCMCglmm](https://CRAN.R-project.org/package=MCMCglmm): `MCMCglmm()`
  - [mixor](https://CRAN.R-project.org/package=mixor): `mixor()`
  - [ordinal](https://CRAN.R-project.org/package=ordinal): `clmm()`, `clmm2()`
